Two Mode Networks and Relational Similarity

### 2023-08-30 main.tex

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.jpg}
    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
    \label{fig:pg-reflections}
\end{figure}
\section{Correspondence Analysis and Dual Centrality}
The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected. In that paper, Bonacich also considered CA in passing but dismissed it as providing a measure of centrality. Instead, Bonacich noted that ``[r]ather than centrality, what they seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.

@@

To see the links between CA, reflective centralities, and clustering in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be seen as giving the similarity between pairs of people weighted by the size of the groups they belong to. The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{Plots/pg-reflections.jpg}
    \caption{Reflective centrality trajectories of persons and groups (even reflections).}
    \label{fig:pg-reflections}
\end{figure}
\section{Correspondence Analysis, Dual Centrality, and Clique Identification}
The reader may ask what the point of going through all these reflective centralities, since \citet{bonacich1991simultaneous} already developed a dual conception of centrality in two-mode networks based on a very similar idea: Define the centralities of entities in one mode based on the centralities of entities in the other mode to which they are connected. In that paper, Bonacich also considered CA in passing but dismissed it as providing a measure of centrality. Instead, Bonacich noted that ``[r]ather than centrality, what they seem to capture is membership in the two cliques that attended two different sets of events'' \citeyearpar[164]{bonacich1991simultaneous}.

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{Plots/ca-v-bon-reord.png}
    \caption{Southern Women Affiliation matrix with re-ordered rows and columns: (a) row and columns re-ordered according to first CA dimension, (b) row and columns re-ordered according to Eigenvector centrality.}
    \label{fig:ca-v-bon}
\end{figure}

Figure~\ref{fig:ca-v-bon} illustrates the differences between $C^R$ and $C^B$. Each panel shows the original Southern Women affiliation matrix but with rows and columns re-ordered according to the first CA dimension (a) and the Bonacich centrality (b). In the plot, a square is colored red if it has a one in the affiliation matrix and is colored white if it has a zero. As we can see, the two matrices have distinct structures, with the $C^R$ re-ordered affiliation matrix having a block-diagonal structure and the $C^B$ re-ordered affiliation matrix having a triangular structure. Accordingly, the $C^R$ centrality rankings reveal a clique-like partition between two groups of women who selectively attend two groups of events (on the top-right and bottom-left of the plot). The traditional eigenvector centrality re-ordering, on the other hand, reveals a classic core-periphery partitioning \citep{borgatti2000models}, with a group of highly active women (on the bottom-right) who selectively co-participate in highly attended events (on the lower right-hand side of the plot) and less active women (in the upper half) who go to less well-attended events (on the right side). The two centralities extract qualitatively different information from the two-mode network, with $C^R$ geared toward clique partitioning and $C^B$ more focused on finding a core of well-connected persons and groups.

To further appreciate the links between CA, reflective centralities, and clique-like clustering in two-mode networks, note that $S_p = ADg^{-1}A^T$ in~\ref{eq:dam1} can be seen as giving the similarity between pairs of people weighted by the size of the groups they belong to. The same interpretation can be given to $S_g = A^TDp^{-1}A$ in~\ref{eq:dam2}.
